---
title: 'Exploration 4: Two Variables Relationships as Causal Relationships'
author: 'Jake Bowers'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
bibliography: classbib.bib
fontsize: 10pt
geometry: margin=1in
mainfont: "Crimson Text"
graphics: yes
output:
  html_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 4
  pdf_document:
    latex_engine: xelatex
    fig_caption: yes
    fig_height: 4
    fig_width: 4
---

<!-- Make this document using library(rmarkdown); render("exploration1.Rmd") -->
\input{mytexsymbols}


```{r include=FALSE, cache=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).
# knitr settings to control how R chunks work.

## To make the html file do
## render("exploration1.Rmd",output_format=html_document(fig_retina=FALSE))
## To make the pdf file do
## render("exploration1.Rmd",output_format=pdf_document())

require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small",    # slightly smaller font for code
  echo=TRUE,
  results='markup',
  strip.white=TRUE,
  fig.path='figs/fig',
  cache=FALSE,
  highlight=TRUE,
  width.cutoff=132,
  size='footnotesize',
  out.width='.9\\textwidth',
  fig.retina=FALSE,
  message=FALSE,
  comment=NA)
```


```{r initialize,echo=FALSE}
##First, just setup the R environment for today:
if(!file.exists('figs')) dir.create('figs')

options(SweaveHooks=list(fig=function(){
			   par(mar=c(3.5, 3, 1.1, 0),
			       pty="s",
			       mgp=c(1.5,0.5,0),
			       oma=c(0,0,0,0))},
			 echo=function(){options(continue=" ") ##Don't show "+" prompts,
			 options(prompt=" ")
			 }),
	digits=4,
	scipen=8,
	width=132
	)
options(error=function(){options(prompt="> ",continue="+ ");NULL})
```


```{r datasetup, cache=TRUE}
download.file("http://jakebowers.org/Data/ANES/anes_pilot_2016_csv.zip",destfile="anespilot2016.csv.zip")
unzip("anespilot2016.csv.zip")
anespi16<-read.csv("anes_pilot_2016.csv",as.is=TRUE,strip.white=TRUE)
```

You friend calls back. "I really appreciated all of the work that you did with
describing the relationships between age and Trump support. Of course, as soon
as your results came in, the team began to bicker: "Older people like Trump
because they are just inherently conservative. It is not age per se that causes
this relationship, but the fact that ideology changes over time within the life
span." then the others argued, "First, there is no fact that ideology changes
over time within the life span. Second, I think that this relationship just
shows that people who are older have had more time to dislike Clinton --- the
young people don't know that much about the Clintons, and so they are not yet
disillusioned." And other group argued, "So one of you is saying that age
causes Trump support because ideology is related to age? And the other is that
age causes Trump support because political knowledge and experience is related
to age? But, both of you seem to be ignoring the nonlinear descriptions that we
did. Didn't it seem like there were different age groups that responded to
Trump differently? Don't you think that this is caused by the fact that the
ethnic composition of the country has changed, and that Trump consistently
offends non-white voters? That is, that the relationship is caused by the
underlying relationship between age and ethnicity and Trump's offensive
comments?"

"I started to try to calm the waters in my team, but I only go this far."

```{r}
table(anespi16$fttrump,useNA="ifany")
table(anespi16$birthyr,useNA="ifany")

anespi16$fttrump[anespi16$fttrump==998]<-NA
anespi16$age <- 2016-anespi16$birthyr

summary(anespi16$age)
summary(anespi16$fttrump)
summary(anespi16$lcself)
anespi16$libcon <- ifelse(anespi16$lcself==8,NA,anespi16$lcself)

res <- sapply(1:7,function(lc){
		      coef(lm(fttrump~age,data=anespi16,subset=libcon==lc))
	})
colnames(res) <- 1:7

library(MASS)
res2 <- sapply(1:7,function(lc){
		      coef(rlm(fttrump~age,data=anespi16,subset=libcon==lc))
	})
colnames(res2) <- 1:7
```

"First, I was getting confused about the use of the word, 'cause'. I didn't quite know what they meant. And, I didn't know how to start using the data on hand to engage with these alternative causal explanations if I didn't know what they meant by cause. The problem is now that I'm back abroad, I can't ask them. What is your best guess? Can you explain it to me? For example, why would thinking about cause one way help me use data to engage with these kinds of arguments? And, of course, I'm relying on you for your advice about which explanation is correct. Please help!  I'm also wondering about your own favorite explanation. One that is not one of those three. What is it? What is the evidence in favor of it or against it?"



# References

