---
title: 'Exploration 2: Description of Relationships'
author: 'Jake Bowers'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
bibliography: classbib.bib
fontsize: 10pt
geometry: margin=1in
mainfont: "Crimson Text"
graphics: yes
output:
  html_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 4
  pdf_document:
    latex_engine: xelatex
    fig_caption: yes
    fig_height: 4
    fig_width: 4
---

<!-- Make this document using library(rmarkdown); render("exploration1.Rmd") -->
\input{mytexsymbols}


```{r include=FALSE, cache=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).
# knitr settings to control how R chunks work.

## To make the html file do
## render("exploration1.Rmd",output_format=html_document(fig_retina=FALSE))
## To make the pdf file do
## render("exploration1.Rmd",output_format=pdf_document())

require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small",    # slightly smaller font for code
  echo=TRUE,
  results='markup',
  strip.white=TRUE,
  fig.path='figs/fig',
  cache=FALSE,
  highlight=TRUE,
  width.cutoff=132,
  size='footnotesize',
  out.width='.9\\textwidth',
  fig.retina=FALSE,
  message=FALSE,
  comment=NA)
```


```{r initialize,echo=FALSE}
##First, just setup the R environment for today:
if(!file.exists('figs')) dir.create('figs')

options(SweaveHooks=list(fig=function(){
			   par(mar=c(3.5, 3, 1.1, 0),
			       pty="s",
			       mgp=c(1.5,0.5,0),
			       oma=c(0,0,0,0))},
			 echo=function(){options(continue=" ") ##Don't show "+" prompts,
			 options(prompt=" ")
			 }),
	digits=4,
	scipen=8,
	width=132
	)
options(error=function(){options(prompt="> ",continue="+ ");NULL})
```


"Thanks!" Your friend's voice sounds much calmer than it did when you last
spoke. "My tasks this week are much less scary than engaging with long-tailed
and zero-inflated variables like hours spent helping others and groups." When
you ask what she is up to, she replies,"I do need some help. At a recent
meeting we had a debate about two different statements. Can you help me by
providing answers based in data? We got into a shouting matching and only
stopped when I mentioned the word 'data'. I don't know much
about data. Can you help? Here are the two statements."

> Support for anti-immigration populists like Trump or the UKIP party arises from anti-feminism more than nativism."

> In countries like the UK, education and income are not strongly related."

"I managed to find a survey that is supposed to be representative of people in the UK."

```{r readdata}
# library(foreign) ## For older Stata files
library(readstata13)
bes<-read.dta13("http://jakebowers.org/Data/BES2015/bes_f2f_original_v3.0.dta",convert.factors=FALSE)
```

"And I found a (codebook)[http://jakebowers.org/Data] too. And have learned enough about R to investigate some variables."

```{r sometabs, results="hide"}
table(bes$r03,useNA="ifany") ##  Feminism
table(bes$Age,useNA="ifany")
table(bes$y01,useNA="ifany") ## Income
table(bes$j05,useNA="ifany") ## Immigration
table(bes$education,useNA="ifany")
table(bes$b02,useNA="ifany")
table(bes$b04,useNA="ifany")
## 1 if voted or would have voted for ukip
bes$ukipvoter <- as.numeric(ifelse(is.na(bes$b02),bes$b04==8,bes$b02==7))
table(bes$ukipvoter,useNA="ifany")
table(bes$d01,useNA="ifany") # Party ID
```

"What do you think? Between these variables, and the others in the codebook,
can you provide some evidence for or against those statements? Since this was
such a heated discussion in our staff meeting, I'd like to see at least three
ways to describe the relationships including that least squares stuff that
people always talk about. And I'll need you to justify the weaknesses and
strengths of the tactics you choose to use. And, especially for the least
squares description, can you please help me understand why it is called least
squares and why I should care about least squares versus most squares or least
something else?  Right now, we just want to describe relationships so
$p$-values or confidence intervals would not be helpful."

```{r}
lm1<-lm(ukipvoter~r03,data=bes,subset=r03!=-1 & j05!=-1)
predlm1 <- predict(lm1,newdata=data.frame(r03=c(1,5)))
predlm1[1] - predlm1[2]

bes$proweq<-(bes$r03-1)/4
bes$proweq[bes$proweq<0] <- NA
table(bes$proweq,bes$r03,useNA="ifany")

lm1a <- lm(ukipvoter ~ proweq , data=bes, subset=j05!=-1)

bes$proimmig <- bes$j05 - 1
with(bes,table(proimmig,j05))
lm2<-lm(ukipvoter~proimmig,data=bes,subset=r03!=-1 & j05!=-1)
predlm2<-predict(lm2,newdata=data.frame(proimmig=c(0,1)))
predlm2[1] - predlm2[2]

lm1b <- lm(ukipvoter ~ proweq + I(proweq^2), data=bes, subset=j05!=-1)

line1 <- .1 - .3 * bes$proweq
diffs1 <- bes$proweq - line1
sum(diffs1^2,na.rm=TRUE)

diffslm1a <- residuals(lm1a)
sum(diffslm1a^2)

myssr<-function(a,b){
  line1 <- a + b * bes$proweq
diffs1 <- bes$proweq - line1
sum(diffs1^2,na.rm=TRUE)
}

as <- runif(100,-1,1) ##.145 + seq(-.1,.1,length=100)
bs <- -.06 + seq(-.1,.1,length=100)
inputs <- expand.grid(as,bs)

res<-matrix(nrow=100*100,ncol=3)
for(i in 1:nrow(res)){
  res[i,1]<-myssr(inputs[i,1],inputs[i,2])
  res[i,2]<-inputs[i,1]
  res[i,3]<-inputs[i,2]
}



```


